Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-04-09 Paper.Rmd
Cedar creek results modified and explained accordingly

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes results for Q1 for the bivariate datasets. For all $4$ 
bivariate datasets, the two variables were always significantly related 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC (highest-AIC-weight) copula 
(Table \ref{tab_bivar_summary}, row 2). 
The N copula was always a poor fit compared to the highest-AIC-weight copula 
(Table \ref{tab_bivar_summary}, rows 3-4), answering the first part of Q1. 
Lower- and upper-tail statistics (Table \ref{tab_bivar_summary}, rows 7-8) 
differed substantially from 0, the tail dependence statistics of 
the N copula, answering the second part of Q1; and also differed substantially from each other
(Table \ref{tab_bivar_summary}, row 9),
answering the third part of Q1. The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse. 
Although these results convincingly show non-normal copula structure, the 
model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula was a poor fit in some cases 
(Table \ref{tab_bivar_summary}, rows 5-6).
But our nonparametric statistics confirmed the same conclusions 
(Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}).
Detailed results for each dataset are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} 
and Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.

>>>>

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes results for Q1 for the bivariate datasets. For all $4$ 
bivariate datasets, the two variables were always significantly related 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC (highest-AIC-weight) copula 
(Table \ref{tab_bivar_summary}, row 2). 
The N copula was always a poor fit compared to the highest-AIC-weight copula 
(Table \ref{tab_bivar_summary}, rows 3-4), answering the first part of Q1. 
Lower- and upper-tail statistics (Table \ref{tab_bivar_summary}, rows 7-8) 
differed substantially from 0, the tail dependence statistics of 
the N copula, answering the second part of Q1; and also differed substantially from each other
(Table \ref{tab_bivar_summary}, row 9),
answering the third part of Q1. The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse. 
Although these results convincingly show non-normal copula structure, the 
model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula was a poor fit in some cases 
(Table \ref{tab_bivar_summary}, rows 5-6).
But our nonparametric statistics confirmed the same conclusions 
(Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}) for first three bivariate datasets.
The fourth bivariate datasets we analyzed was above-ground biomass and percent cover data (i.e. 
the data used to compute Shannon's diversity index) for the year 2000, but we did the same analysis 
with Cedar creek data sampled over 168 plots from other available years (1996-1999, 2007) too. For 1996 and 
2007 copula structures showed no dependence between two variables concerned, for other years (1997-1999), 
we found similar results as of year 2000 (i.e. best fit non-Normal copula and overall lower tail-dependence).
Due to smaller sample size relative to other three bivariate datasets, we chose a wider range with 
non-parametric analysis (see Appendix \ref{SI-Q1bivarnonparam}). Detailed results for each dataset are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.

### 2019-04-09 SupportingInformation.Rmd
Cedar creek results modified and explained accordingly

#----------------------------------------------------------------------------------------------
# This sections selects 168 plots considered for both biomass and percent cover datasets in Tilman's papers 2001, 2006
i<-5#for Yr=2000
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
#------------------------------------------------------
id_plot_nosp<-which(db$NumSp==0 & db$Biomass==0) # These plots have no sp. seeded initially and have no biomass data
plots_no_biomass<-sort(unique(db$Plot[id_plot_nosp]))
plots_no_biomass

plots<-sort(intersect(db$Plot,dp$Plot)) # common plots for both datasets for a specific year
length(plots)

excluded_plots<-intersect(plots,plots_no_biomass)

plots_168<-setdiff(plots,excluded_plots)

#--------------------------------------------------------------------------------------------
pdf("./Results/Ceder_creek_results/raw_HB_all_yr.pdf",height = 6,width = 10)
op<-par(mfrow=c(2,3),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
for(i in 1:length(yr)){
  db<-subset(r_ple,Year==yr[i])
  dp<-subset(r_pce_clean,Year==yr[i])

@@

For all four bivariate datasets the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of the real-data value in 
the distribution of surrogate values was computed. If ranks were $>975$ or $<25$, asymmetry of tail 
dependence was significant compared to a N-copula null hypothesis. Ranks were 1, 1, 1, 998
<!--***DAN: Shya, please make the quantities in the previous line automatic links-->
for, respectively, the soil C and N data, bird masses and metabolic rates, mammal masses
and metabolic rates, and the Cedar Creek data. Thus the first three datasets showed stronger 
upper- than lower-tail dependence, and the Cedar Creek data showed the reverse, confirming the 
results based on model selection presented in the main text.

@@

for(i in 1:length(stat_cc_ga)){
  temp<-bivfunctionplot(v=cop_HB_all_yr[[i]],
                             resloc="./Results/Ceder_creek_results/",
                             nametag=paste0(names(cop_HB_all_yr)[i],"cc_numbin_10",sep=""),
                             numbin = 10)
  stat_cc_ga[[i]]<-temp
}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data. Same format as Fig. \ref{fig_soilCN_nonparam}, see caption of that figure.\label{fig_cc2000_nonparam}}
\end{center}
\end{figure}

<!--Displaying a table with npa summary results for Cedar creek data-->
```{r tab_cc_asym_2000, echo=F, results="markup"}
tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
#tab_cc_asym_2000[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_cc_asym_2000,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
             caption = "Cedar Creek dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_cc_asym_2000}",
             booktabs=T)
```

>>>>

#----------------------------------------------------------------------------------------------
#=======================================================
# This sections selects 168 plots considered for both biomass and percent cover datasets in Tilman's papers 2001, 2006
plot_info<-read.delim("./Data/Ceder_creek_e120/edited_plots.txt")

id_excluded_plots<-which(plot_info$Species=="Bargr") # there were 4 plots with no sp. seeded

plots_168<-plot_info$Plot[-id_excluded_plots]

#------------------------------------------------------
# You can check that here too
s<-subset(r_ple,NumSp==0)
sb<-sort(unique(s$Plot))

s<-subset(r_pce_clean,NumSp==0)
spc<-sort(unique(s$Plot))

sb==spc # This should be all TRUE for 4 not seeded plots with id : 126, 166, 244, 248
#=======================================================
#--------------------------------------------------------------------------------------------
pdf("./Results/Ceder_creek_results/raw_HB_all_yr.pdf",height = 6,width = 10)
op<-par(mfrow=c(2,3),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
for(i in 1:length(yr)){
  db<-subset(r_ple,Year==yr[i])
  dp<-subset(r_pce_clean,Year==yr[i])

@@

@@

for(i in 1:length(stat_cc_ga)){
  temp<-bivfunctionplot(v=cop_HB_all_yr[[i]],
                             resloc="./Results/Ceder_creek_results/",
                             nametag=paste0(names(cop_HB_all_yr)[i],"cc_numbin_4",sep=""),
                             numbin = 4)
  stat_cc_ga[[i]]<-temp
}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_4_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$. If the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.\label{fig_cc2000_nonparam}}
\end{center}
\end{figure}

<!--Displaying a table with npa summary results for Cedar creek data-->
```{r tab_cc_asym_2000, echo=F, results="markup"}
tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
#tab_cc_asym_2000[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_cc_asym_2000,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
             caption = "Cedar Creek dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.25}-\\cor_{0.75,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.25}-\\Ps_{0.75,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.75,1}-\\Dsq_{0,0.25}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons. A table entry $>X$ indicates the value of the given statistic on the data was higher than its value on $X$ of the surrogates, so entries of the form $>X$ for $X$ equal to $975$ or above would indicate that lower-tail dependence was significantly stronger than upper-tail dependence.\\label{tab_cc_asym_2000}",
             booktabs=T)
```

For first three bivariate datasets in Table \ref{MT-tab_bivar_summary},
the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
for 1000 Kendall-preserving surrogate datasets with N copulas. Due to relatively smaller sample size,
for Cedar creek data we used a wider range of boundaries $(l_b,u_b)$ to compute $P_{0,0.25}-P_{0.75,1}$.
For each data, rank of the said statistics with real-data value in 
the distribution of surrogate values was computed. 
If ranks were $>975$ or $<25$, asymmetry of tail 
dependence was significant compared to a N-copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
for, respectively, the soil C and N data, bird masses and metabolic rates, mammal masses
and metabolic rates, and the Cedar Creek data. Thus the first three datasets showed stronger 
upper- than lower-tail dependence, and the Cedar Creek data which showed the reverse though not significantly with
non-parametric analysis, confirming the 
results based on model selection presented in the main text (see Table \ref{MT-tab_bivar_summary}, row 9 
showed Cedar creek had relatively weak tail-dependence than the other three datasets).
